\documentclass[12pt,titlepage,letter]{article}
\usepackage[margin=1in]{geometry}

\title{Bylaws of the\\Davis Sailing Team\\at\\the University of California, Davis}
\date{}

\begin{document}

\maketitle

\tableofcontents

\section{Name}

The name of this organization shall be the Davis Sailing Team at the University of California, Davis and shall be referred to herein as ``the team".

\section{Purpose}

The team will be organized and operated exclusively for educational purposes within the meaning of Section 501(c)(3) of the Internal Revenue Code of 1954, as amended, and its activities shall meet the following purposes:

\begin{enumerate}
\item
The instruction, training, and education of students and affiliated members at the University of California, Davis in seamanship, safety, and the skills and techniques of sailboat racing as sanctioned by the Inter-Collegiate Sailing Association of North America, Inc. (here after referred to as ``ICSA") and by the Pacific Coast Collegiate Sailing Conference (here after referred to as ``PCCSC").
\item
The sponsorship, conduct, supervision, development and provision for sailboat racing competition among students and affiliated members at the University of California, Davis for the purposes of improving and developing the sailboat racing techniques.
\item
The improvement, advancement, and encouragement of sailboat racing at the University of California, Davis and the development of the high ideals and sportsmanship preeminent in sailboat racing competition.
\item
By means of the foregoing activities, the identifying, developing, and training of potential representatives for US SAILING events, Olympic, Pan American, and other international competition.
\end{enumerate}

The team shall be empowered to own, hold, operate, and administer real and personal property solely in furtherance of the previously stated educational purposes and to solicit, accept, hold, and administer contributions howsoever received; whether by deed, gift, bequest, device or otherwise. The team shall be empowered to do any and all lawful acts and things and to engage in any and all lawful activities which may be necessary for or convenient to the carrying out of the foregoing activities, but solely in the furtherance of said educational purposes.  

No part of the net earnings of the team shall inure to the benefit of any individual and no part of the activities of this team shall be carrying on propaganda or otherwise attempting to influence legislation, nor shall this team participate in or intervene in (by publishing or distributing statements or otherwise) any political campaign on behalf of any candidate for public office.  No officer, member, or affiliate of this team shall receive or be entitled to receive any pecuniary profit from the operations of this team except reasonable compensation for services rendered or expenses incurred. 

This team shall not engage in any prohibited transaction described in Section 503 of said Internal Revenue Code.  
In the event of the dissolution of this team, the assets of the team then remaining will be distributed as specified by section \ref{Dissolution} provided the successor(s) qualify under Section 501(c)(3) of said Internal Revenue Code, otherwise to those members of the team that are qualified under said Section 503(c)(3), and if none of the foregoing are qualified under said Section 503(c)(3) then to one or more other organizations whether or not (in each instance) incorporated, organized, and operated exclusively for one or more purpose, each of which is dominated and defined in said Section 501(c)(3).

\section{Membership \label{member}}
\subsection{Eligibility \label{member:eligibility}}
Any person is eligible to become a member of the team provided they meet the following requirements:
\begin{enumerate}
\item
A person must be enrolled in the University of California, Davis as either an undergraduate or graduate student, and be in good standing as determined by the University of California, Davis.
\item
A person must be a member of good standing of the Lake Washington Sailing Club.
\item
A person must comply with all the articles of the By-Laws, complete any associated membership forms, and pay any necessary dues.
\end{enumerate}
Consideration for membership cannot be denied on the basis of gender, race, religion, creed, national origin, sexual orientation, disability, marital status, or experience.  

\subsection{Member's Rights and Duties \label{member:duties}}
Once a person has met the stated requirements they shall become a member of the team and be entitled to and accountable for the following:
\begin{enumerate}
\item
The member must pay all dues and fees levied by the team. 
\item
The member is entitled to participate in events, such as regattas and practices.
\item
The member is entitled to vote.
\item
The member is entitled to membership of affiliated organizations, such as the Inter-Collegiate Sailing Association, provided they meet the affiliated organization's membership requirements.
\item
The team member shall conduct themselves in a safe and reasonable manner at any team event.
\end{enumerate}

\subsection{Amateur Status \label{member:amateur}}
At no time shall any member receive compensation for the act of sailing or conduct business for the team unless they are performing their duties as a team officer or authorized by the Executive Committee (see Article \ref{executive}).

\subsection{General Membership \label{member:general}}
The entire membership shall be referred to as the general membership. The general membership can make decisions on issues put forward by the Executive Committee (see section \ref{executive}) or a function of the By-Laws, by a majority vote of the members present at a general meeting (see Article \ref{meeting:general}). 

\subsection{Term \label{member:term}}
A membership shall be valid from the beginning to the end of one quarter as specified by the academic calendar of the University of California, Davis.

\subsection{Misconduct \label{member:misconduct}}
If a team member conducts themselves in a manner contradictory to any team rules or regulations, the member may be placed on probation, may be suspened, or be expelled from the team. Probation shall consist of a mark on the member's record and the member shall suffer no ill effects but will serve as a warning to the member. For more serious offenses, a member can be suspended from participating in team events for the remainder of their membership term. For severe offenses, the member can be expelled from the team for one year past their violation.  For failing to pay any fee, a member will be automatically suspended and will follow the protocol specified in Section \ref{member:misconduct:automatic}. For conducting themselves in any other manner contradictory to team rules, they will be reviewed based on the protocol specified in Section \ref{member:misconduct:general}.

\subsubsection{Automatic and Indefinite Suspension \label{member:misconduct:automatic}}
If a member fails to pay a fee levied by the team, the member will be placed on automatic suspension. This suspension will be effective until the member pays the fee even if their membership term expires. This automatic suspension will follow the following protocol:
\begin{enumerate}
\item
The member must be notified in writing within three days of their violation.
\item
The member has two weeks from the notification to contest the suspension. The member must submit their reasons for contesting the suspension in writing or orally to the Executive Committee (see section \ref{executive}). The Executive Committee will then review and vote by majority to reverse or alter the suspension at the next available Executive Committee meeting (see section \ref{meeting:executive}).
\end{enumerate}

\subsubsection{General Misconduct \label{member:misconduct:general}}
Should a member conduct themselves in a manner contradictory to team rules, a written description of the violation can be submitted to the Executive Committee (see section \ref{executive}) by any person, be it member, non-member or officer. The Executive Committee will then review this description and then vote by majority to accept the violation. Should the Executive Committee vote to accept the violation the following protocol will be followed.
\begin{enumerate}
\item
A mutually agreeable time for both the Executive Committee and the member in question to meet is arranged.
\item
At this meeting the member and their alleged violation will be reviewed by the Executive Committee. For part of the review, the member will be able to provide defense of their actions. The Executive Committee will then vote whether to reprimand the member. For minor first time offenses the member should be placed on probation for the remainder of the term. For moderate, repeated, or if a member is already on probation, the member should be suspended from participating in team events for the remainder of their membership term. For severe offenses the member should be expelled from the team. Should the member accept guilt of their actions, the term of their reprimand will be decided by the Executive Committee provided the reprimand length is less than or equal to their membership term.
\item
Should a member be placed on suspension or be expelled from the team, the general membership must be notified of this decision at the next general meeting (see section \ref{meeting:general}). A member of general membership may motion to contest this decision provided they are seconded. If a motion occurs, the member will be reviewed by the general membership at this meeting. During this review the member will provide a defense of their actions. Once the review is complete the general membership shall then vote by majority to uphold the decision of the Executive Committee.
\end{enumerate}

\section{Officers and Executive Committee \label{executive}}
\subsection{Purpose}
The team shall be directed by a committee of five elected officers that will make decisions based on a majority vote. This committee will be referred to as the Executive Committee and will consist of five officers: Executive Officer, Executive Sailing Officer, Vice President of Finances, Vice President of Records and Vice President of Public Relations. Each of these officers must be a member in good standing and will be elected by the general membership as prescribed the election process in Section \ref{elections}, provided they have been a member for one membership term.
\subsection{Duties}
The duties of each officer are as follows and they will perform these duties along with the general duties. Every officer will be responsible for performing their duties to the best of their abilities in a timely manner.
\begin{description}
\item[Executive Officer] shall preside over general membership and Executive Committee meetings (see section \ref{meeting}) of the team, oversee the organization of the team, oversee the planning of team events including the team calendar (with the exception of practicing and racing), oversee the finances of the team, oversee the development of the team and enforce the constitution and By-Laws of the team. The Executive Officer shall have the power to change, cancel, or otherwise alter meeting times or locations provided the general team membership is aware of the change two days prior to the new time or location. The Executive Officer will also be responsible for producing an agenda for the meetings.
\item[Executive Sailing Officer] shall be responsible for the sailing aspects of the team including but not limited to the responsibility of the fleet, the organization and oversight of practices and the communication with the appropriate sailing administrations. The Executive Sailing Officer shall have the power to change, cancel, or otherwise alter practice times or locations provided the general membership was aware of the change two days prior to the new time or location.
\item[Vice President of Finances] shall be responsible for financial matters of the team including but not limited to the collection of dues, management of all team accounts, and providing a team budget.
\item[Vice President of Records] shall be responsible for all inter-team communications. The Vice President of Records shall provide minutes for meetings and a contact directory of all active members. The Vice President of Records shall also be responsible for the creation and the distribution of the newsletter.
\item[Vice President of Public Relations] shall be to oversee outside contributions to the team. The Vice President of Public Relations shall also be responsible for acquiring the team's apparel, recruiting new members and social aspects of the team.
\end{description}

\subsection{Term}
All officers shall hold office for one year. Their term shall commence and end at the end of spring quarter as specified by the calendar at the University of California, Davis. Officers elected before their term commences will still be required to attend Executive Committee meetings, see section \ref{meeting:executive}.

\subsection{Resignation}
Any officer may resign by notifying the Executive Officer in writing. Should the Executive Officer wish to resign, the written resignation will be delivered to the Executive Sailing Officer. Such resignation shall be accepted in a timely manner, provided any dues and all other indebtedness to the team have been paid.  The officers, by a majority vote, may waive the payment of such dues and other indebtedness owed to the team as a condition of accepting the resignation.

\subsection{Misconduct}
Should an officer fail to perform any duties or act in a manner contradictory to team rules or regulations, a written grievance may be delivered to the Executive Committee by any member of the team. The Executive Committee shall follow the same protocol as a membership violation, except the officer in question may not vote. The Executive Committee can remove an officer from their office by a majority vote. Should any misconduct vote end in a tie, the vote will then pass to the general membership which can reprimand the officer based on a two-thirds majority vote of the general membership.

\subsection{Volunteer Status}
At no time shall an officer receive payment for conducting their duties. However, an officer may be authorized by an Executive Committee decision to conduct business on the team's behalf. Any cost incurred to the officer for conducting the business shall be reimbursed provided the officer presents receipt.

\section{Dues}
\subsection{Membership Dues}
	Dues will be collected upon commencement of membership. Dues will not be refunded under any circumstances and should a member withdraw from the team, they shall be responsible for paying their dues and any other debts owed. Dues can be increased only by amending the Bylaws (see section \ref{amendments}) and will go into effect at the beginning of the following term. The amount of the dues shall be set for twenty-five dollars (\$25) per member each term.

\subsection{Regatta/Event Fee}
A member that participates in a regatta or event that incurs an expense to the team will be responsible for paying a fee for that regatta or event. This fee and collection date will be determined by the Executive Committee and made known to the participating members at least three days prior to the regatta or event. Should a member fail to pay the fee by the collection date, an additional late fee will be levied. Should the cost of the regatta or event exceed the total fee collected by the team, the difference will be paid by the treasury, not by the participating members. Should the cost of the event or regatta be less than the total fee collected by the team, the difference shall be added to the treasury, not refunded to the participating members.


\section{Meetings \label{meeting}}
Meetings shall be divided into general meetings, Executive Committee meetings and practices.
\subsection{General Meetings \label{meeting:general}}
General meetings shall be held on a weekly basis for the benefit of informing the team. These meetings will be open to the public and any change to the time or meeting place must be announced three days prior to the meeting. A minimum of three Executive Committee officers must be present at the meeting and one must produce minutes for the meeting. Minutes from the previous meeting must be displayed in a public area and need only be read upon request. The general meetings will follow the following format:
\begin{enumerate}
\item
The Executive Officer or other presiding officer calls the meeting to order.
\item
The Vice President of Records asks for a motion and a second to read the minutes from the previous meeting.
\item
Informational items are reported.
\item
Any items that require a decision to be made will first be discussed. Once the discussion is complete, motions will be accepted provided they are seconded. Once a motion is accepted, the members present will then vote by majority to pass the decision.
\item
The officers present make their reports.
\item
Any members present make any reports.
\item
Any public comment is made.
\item
The floor is opened to general discussion.
\item
The meeting is adjourned.
\end{enumerate}

\subsection{Executive Committee Meetings \label{meeting:executive}}
Prior to a general meeting the Executive Committee must hold a meeting to discuss the general management of the team. Minutes for this Executive Committee meeting must be produced. The general presentation for a general meeting of Executive Committee meeting is as follows:
\begin{enumerate}
\item
The Executive Officer calls the meeting to order.
\item
The Vice President of Records reads the minutes from last meeting.
\item
Each officer gives a report on the duties they have been performing.
\item
Any informational items are discussed.
\item
Any items that require a decision to be made will first be discussed. Once the discussion is complete, motions will be accepted provided they are seconded. Once a motion is accepted, the Executive Committee will then vote by majority to pass the decision.
\item
Once all reports are finished any member present may make a report and then the floor will be open for general discussion.
\end{enumerate}
Should the Executive Committee be unable to pass a vote due to a tie; the vote will pass to the General Membership which will vote to pass the decision by a majority.

\subsection{Practices \label{practices}}
Practices shall be held at least once a week and shall involve the sport of sailing. Should the team be participating in a regatta or other event, practice need not be held. Practice may also be canceled by the Executive Sailing Officer as specified in section \ref{member:duties}. Any member in good standing with the team shall be eligible to attend a practice.

\section{Elections \label{elections}}

\subsection{Date}
Elections for officer positions shall be held annually during spring quarter as specified by the University of California, Davis academic calendar. Elections should take place before the end of the quarter allowing time for the incoming board to transition. 

\subsection{Voting Members}
Any member in good standing shall be allowed to vote provided they attend the election day. Absent members will not count in the election nor may they have another member vote in their stead. A voting member may also be a candidate in the election but they may not to vote for themselves. Members that are also candidates may vote in the election provided they do not vote for the position they are running for.

\subsection{Election Format}
The election shall be held at a general meeting. Before the election occurs a proctor will be selected by the current executive committee. This proctor shall not be a candidate for any position nor may they vote in the election. The election will then begin and adhere to the following format:

\begin{enumerate}
\item
	Nominations for the position will be accepted.
\item
	Each Candidate will present an argument for why they would like to hold the office.
\item
	The voting members will discuss the candidate's platform and qualities with the candidate present.
\item
	The voting members will discuss the candidate's platform and qualities without the candidate present.
\item
	Members shall then cast their votes by means of secret ballot.
\end{enumerate}
Should a tie occur the process will be repeated between the two tied candidates until a winner is decided.

An election may be contested by any member of the team on the following grounds:
\begin{enumerate}
\item
Ballots were miscounted.
\item
The proctor influenced the election.
\item
A member that wished to be a voting member was unable to vote.
\item
A member that wished to be informed of an argument was denied.
\item
Not all candidates were able to present arguments in the same format.
\end{enumerate}
The former Executive Committee will rule on the validity of the contested election and will render the sole judgment based on the previous criteria. If the election is not valid, a new election will be held with a new proctor, and the former proctor shall not be able to vote in the new election. Should the former proctor be a member of the former Executive committee, they shall be barred from the decision to approve the election.

\section{Amendments \label{amendments}}
With the exception to section \ref{meeting} all amendments shall follow the protocol specified by section \ref{amend:general}. Any amendments to section \ref{meeting} and its subsections will follow the protocol specified by section \ref{amend:meetings} and its subsections.

\subsection{General Amendments \label{amend:general}}
Amendments to the Bylaws must be submitted in writing before a general meeting. The present membership will vote by majority to pass the amendment. Should the amendment pass the Executive Committee shall vote by majority to accept the amendment. Should the amendment be accepted it will amended to the Bylaws. Should the Executive Committee reject the amendment, must be altered to repeat the process.

\subsection{Amendments to Section \ref{meeting} \label{amend:meetings}}
With the exception of section \ref{practices}, the Executive Officer may amend or suspend any aspects of section \ref{meeting:general} or \ref{meeting:executive} provided that this  amendment or suspension be less than one membership term. For permanent amendments to section \ref{meeting:general} the amendment must follow the protocol specified in section \ref{amend:general}. For permanent amendments to section \ref{meeting:executive}, the amendment must be passed by a majority vote of the Executive Committee.

The Executive Sailing Officer may amend or suspend any aspect of section \ref{practices} provided it was less than one membership term. For permanent amendments to section \ref{practices} must follow the protocol specified in section \ref{amend:general}.

\section{Dissolution \label{Dissolution}}
Should the team dissociate, any remaining assets will be presented to the Pacific Coast Collegiate Sailing Conference provided they are exempt under Section 501c(3) at the time the dissolution takes place. Should they be ineligible to receive the assets, the assets will be presented to University of California, Davis or the City of West Sacramento for a public purpose.

\section{Ratification}
10/5/2008

\pagebreak

\section{Amendments}
\subsection{Financial Aid}
Any member in good standing may request funding for any sailing related activity provided the activity will be mutually beneficial to both the team and the member. The member must submit a request to the Executive Committee in writing with a detailed expiation for the need and amount of funding. The Executive Committee will award funding as they see fit and is not obligated to meet the request amount. Aid will be awarded on first come first serve basis. No aid will be awarded if the total funds of the team are less than \$1,000.

Ratified 10/27/2011

\subsection{Amendments to Section \ref{executive}}
The following amendments expand and clarity section 4.
\subsubsection{Officer Duties}
\begin{description}
\item[Executive Officer] will be given access to any financial accounts. The executive officer may not vote in any motions put forward to the Executive Committee. However the executive officer has the power to veto any motion made by the Executive Committee. A vetoed motion may be modified and re-motioned by the Executive Committee. Should the modified motion be vetoed as well, the motion will pass to the general membership.
\item[Executive Sailing Officer] will be given access to any financial accounts.
\item[Vice President of Finances] will be responsible for filling out and completing IRS Form 990 annually by December 31st.
\end{description}

Ratified 10/27/2011  

\subsection{Absent officer protocol}
Should an officer be absent from a meeting, a prearranged replacement may be given the powers of the officer provided they were appointed by the absent officer prior to the meeting.
  
Should an officer be participating in a planned academic leave of absence, study abroad, or other related program for one quarter or less (as defined by the University of California, Davis), the officer may recommend a permanent replacement to hold the office while they are participating in their activity. This permanent replacement must be a member in good standing and will have all the powers and responsibility of the position. The Executive Committee will vote to accept the replacement officer. Should the vote fail, a new candidate will be recommended by the absent officer. 
  
Ratified 10/27/2011  

\subsection{Amendments to \ref{executive}: Watch Officer}

Each event will have one officer responsible for its entire aspects.  This officer will be predetermined by a majority vote of the executive committee. This officer will have the sole responsibility for this event, including but not limited to attendance, transportation, cancellation, etcetera. This officer’s authority will supersede the authority granted in Section 4.2. Should no officer be elected for the event, responsibility will follow the protocol in Section 4.2.

Ratified 5/16/2012

\subsection{Amendments to Section \ref{member}}

\subsubsection{Amendment to Section \ref{member:duties}: Mandatory Eligibility Disclosure}

In addition to rights and duties listed in Section \ref{member:duties} members must affirm their good standing as a student at the University of California, Davis on a quarterly basis. Should a member fail to be in good standing, they will be automatically placed on probation. Should an officer fail to be a student in good standing at the University of California, Davis, they must resign as officer for the remainder of their officer term.

Should a member falsify their standing as a student at the University of California, Davis they will be expelled from the team following the protocol specified in Section \ref{member:misconduct:general}.

Ratified 5/16/2012

\subsubsection{Amendment to Section \ref{member:misconduct}: Definition of Probation}

A member placed on probation will be barred from participating in ICSA and PCCSC sanctioned events. However, a member on probation will be allowed to participate non-sanctioned events, including but not limited to practices, meetings, and social events provided they receive approval by the executive committee prior to the event.

Ratified 5/16/2012

\end{document}
